---
title: "Task 5: Stacked Bar Charts"
author: "Mileva Van Tuyl"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60))
```

# Load Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

# Read Data
```{r cache=TRUE, warning = FALSE, message = FALSE}
GSS2018 <- read_csv(file="./data/GSS2018.csv")
dim(GSS2018)
```

# Process Data
***Fatigue and workstatus from the 2018 GSS Data***    
Respondents were asked the following two questions:  

1. “In the past seven days, how would you rate your fatigue on average?”  
2. "Last week were you working full time, part time, going to   
school, keeping house, or what?"

```{r}
# Drop NA values in the fatigue column 
GSS2018.clean <- GSS2018 %>% drop_na(fatigue) %>% filter(wrkstat != 9)
GSS2018.clean
```

```{r}
# Recode the fatigue and workstatus columns
GSS2018.coded <- GSS2018.clean %>% 
  select(fatigue, wrkstat) %>% 
  transmute(fatigue = fct_recode(as.factor(GSS2018.clean$fatigue), 
                                 None='1', Mild='2', Moderate='3', 
                                 Severe='4', VerySevere='5')) %>%
  mutate(workstatus = fct_recode(as.factor(GSS2018.clean$wrkstat), 
                                 FullTime='1', PartTime='2', TempNotWorking='3', 
                                 UnemployedLaidOff='4', Retired='5', School='6', 
                                 House='7', Other='8'))

GSS2018.coded
```
# Initial Exploratory Data Analysis
Initial analysis to understand how fatigue breaks down by work status.

```{r}
# Compute conditional proportions (i.e. workstatus given a certain level of fatigue)
dataplot <- GSS2018.coded %>% 
  group_by(fatigue) %>% 
  count(workstatus) %>% 
  mutate(prop = n/sum(n)) %>%
  select(fatigue, workstatus, prop)

dataplot
```

```{r}
# Stacked bar chart of workstatus given each level of fatigue)

ggplot(data=dataplot, aes(x=fatigue, y=prop, fill=fct_rev(workstatus))) + 
  geom_bar(stat='identity', position='stack')
```

# 
  

```{r}
# Recode workstatus variable
workstatus.recoded <- GSS2018.clean$wrkstat %>% 
  as.factor() %>% 
  fct_recode(FullTime='1', PartTime='2', TempNotWorking='3', 
             UnemployedLaidOff='4', Retired='5', School='6', House='7', Other='8') %>%
  fct_collapse(Other=c('Other', 'TempNotWorking'), Employed=c('FullTime', 'PartTime')) 

unique(workstatus.recoded)
```


```{r}
# Recode fatigue variable
fatigue.recoded <- GSS2018.clean$fatigue %>%
  as.factor() %>%
  fct_recode(None='1', Mild='2', Moderate='3', Severe='4', VerySevere='5') 

unique(fatigue.recoded)
```

```{r}
# Recode GSS2018 dataset
GSS2018.recoded <- GSS2018.clean %>% 
  select(fatigue, wrkstat) %>% 
  transmute(fatigue = fatigue.recoded) %>%
  mutate(workstatus = workstatus.recoded) %>%
  filter(workstatus != "Other")

GSS2018.recoded
```

```{r}
# Compute conditional proportions (i.e. workstatus given a certain level of fatigue)
dataplot <- GSS2018.recoded %>% 
  group_by(fatigue) %>% 
  count(workstatus) %>% 
  mutate(percentage = 100 * n/sum(n)) %>%
  select(fatigue, workstatus, percentage)

dataplot
```
# Stacked Bar Chart Visualization

```{r}
ggplot(data=dataplot, aes(x=fatigue, y=percentage, fill=fct_rev(workstatus))) + 
  geom_bar(stat='identity', position='stack') +
  scale_x_discrete(limits=c('None', 'Mild', 'Moderate', 'Severe', 'VerySevere'),
                   labels=c('No \nFatigue', 'Mild \nFatigue', 'Moderate \nFatigue',
                            'Severe \nFatigue', 'Very Severe \n Fatigue')) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_discrete(limits=rev(c("Employed", "UnemployedLaidOff", 
                                    "Retired", "School", "House")),
                      labels=rev(c("Employed", "Unemployed", "Retired", "School", 
                                   "House"))) + 
  labs(x = "",
       y = 'Percentage (%)',
       title = 'Associations Between Fatigue and Work Status',
       subtitle = '2018 GSS Data', 
       fill = "Work Status") +
    theme_classic() 

```