---
title: "Gun Control"
author: "Mileva Van Tuyl"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r}
library(naniar)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(vcd)
```

# Read data
```{r}
GSS2016 <- read.csv('./data/GSS2016.csv')
dim(GSS2016)
```

# Processing and Analysis
An analysis of gun ownership and support for background checks when purchasing guns in response to Hillary Clinton's faulty venn diagram from 2016 on gun control (https://www.washingtonpost.com/news/the-fix/wp/2016/05/20/we-fixed-hillary-clintons-terrible-venn-diagram-on-gun-control/). 

The 2016 General Social Survey contains data on gun control opinions in America. 
1. Variable: owngun. Question phrasing: "Do you happen to have in your home (or garage) any guns or 
revolvers?"
2. Variable: gunlaw. Question phrasing: "Would you favor or oppose a law which would require a person 
to obtain a police permit before he or she could buy a gun?"

## Data Processing
First, we must process the data (e.g. recode the responses to the owngun and gunlaw variables and handle missing values).
```{r}
na_strings = c('0', '8', '9')

# Recode the owngun variable
owngun.recoded <- GSS2016$owngun %>%
    as.factor %>%
    fct_recode(Yes = '1', No = '2', Refused = '3')

# Recode the gunlaw variable
gunlaw.recoded <- GSS2016$gunlaw %>% 
  as.factor %>%
  fct_recode(Favor = '1', Oppose = '2')

# Recoded GSS data
GSS2016.recoded <- GSS2016 %>%
    select(owngun, gunlaw) %>%
    replace_with_na_all(condition = ~.x %in% na_strings) %>%
    mutate(owngun = owngun.recoded) %>%
    mutate(gunlaw = gunlaw.recoded)

GSS2016.recoded
```
## Pie Chart
Second, what percentage of respondents favor gun laws? What percentage oppose them?  

```{r}
# Compute the percentage of those in favor and in opposition

data.pie <- GSS2016.recoded %>% 
  drop_na(owngun, gunlaw) %>%
  add_count(gunlaw) %>%
  mutate(percentage = n/ sum(unique(n))) 

data.pie
```
```{r}
# Pie chart 

mycols = brewer.pal(2, 'Set2')

ggplot(data = data.pie, aes(x = "", y = percentage, fill = gunlaw)) + 
  geom_bar(stat='identity') + 
  coord_polar("y", start=0) +
  labs(title = 'Opinions on Gun Control in America',
       subtitle = 'GSS 2016 Data',
       fill = 'Favor or Oppose Gun Permits') + 
  scale_fill_manual(values = mycols) +
  theme_void() + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

# Stacked bar chart
Third, compute the proportion of those in favor and in opposition to gun laws conditioning on whether the respondent owns a gun. 

```{r}
# Computations for conditional probabilities

data.bar <- GSS2016.recoded %>%
    drop_na(owngun, gunlaw) %>%
    filter(owngun != 'Refused') %>%
    group_by(owngun) %>%
    count(gunlaw) %>%
    mutate(prop = 100 * n/ sum(n)) %>%
    select(owngun, gunlaw, prop)

data.bar
```

```{r}
# Bar chart of conditional probabiliies

ggplot(data = data.bar, aes(x = owngun, y = prop, fill = fct_rev(gunlaw))) + 
  geom_bar(stat='identity', position='stack') +
  scale_x_discrete(limits=c('No', 'Yes')) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(title = "Opinions on Gun Control in America", 
       subtitle = 'General Social Survey Data', 
       x = 'Gun Owner',
       y = 'Percentage (%)', 
       fill = 'Favor or Oppose Gun Permits') + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))
```

# Mosaic Plot
```{r}
data.mosaic <- GSS2016.recoded %>%
  filter(owngun != 'Refused') %>%
  mutate(owngun = fct_rev(owngun)) %>%
  droplevels()
  
mosaic(gunlaw ~ owngun, data = data.mosaic, 
       main = 'Opinions on Gun Control in America', 
       labeling_args = list(set_varnames = c(owngun = 'Possession of Gun', 
                                             gunlaw = 'Favor or Oppose Gun Permits')))
```

